\textbf{Teaching a machine to generate an artificial universe}

### 2020-10-14 paper.tex
initial draft

\begin{figure}
\includegraphics[width=\columnwidth]{figures/diagrams/ACGAN_diagram_box.png}
\centering
\caption{Schematic of the GAN framework; including discriminator and generator networks, the latent-space vector, dataset 
         of samples and the backpropagation feedback to model parameters. \textit{Source} and \textit{Class} refer to the
         values of $p_\mathbf{x}$ and $p_\mathbf{c}$ designated by the discriminator on a sample $\mathbf{x}$.}
\label{fig:GAN_diagram}
\end{figure}

@@

% WHAT ARE D AND G? HOW DO D AND G WORK?
The discriminator and generator models can be represented as functions $D$ and $G$ respectively because of their 
mappings as neural networks and parameter dependencies. The discriminator $D$ has image input $\mathbf{x}$ and parameters 
$\bm{\theta}^{(D)}$. The generator $G$ takes a latent space vector $\mathbf{z}$ as input and uses parameters 
$\bm{\theta}^{(G)}$. Information flows through the network and its parameters to create the outputs $D(\mathbf{x})$ or 
$D(G(\mathbf{z}))$ for the discriminator and $G(\mathbf{z})$ for the generator. This is the forward-propagation from 
an input to an output. The model performances are calculated with the objective functions $J^{(D)}$ or $J^{(G)}$.

% HOW DO D AND G WORK TOGETHER IN A GAN AND PLAY THE GAME?
In the case of a GAN, the optimisation landscape is more complex because of the simultaneous learning of the two 
neural networks that both constantly adjust their parameters. The discriminator and generator models in a GAN depend 
on parameters $\bm{\theta}^{(D)}$ and $\bm{\theta}^{(G)}$ respectively. The models typically consist of multi-layer 
perceptrons with the potential for other parametric components such as convolutional layers and the parameters are 
the state of these components at an epoch of training. The generator $G$ draws noise vectors $\mathbf{z}$ from a prior 
distribution $p_{\mathbf{z}}$. The vectors $\mathbf{z}$ are passed through the generator architecture to create a 
sample $\mathbf{x}=G(\mathbf{z};\bm{\theta}^{(G)})$ that is now a sample $\mathbf{x}$ from the model distribution 
$p_{model}$. The model distribution is the distribution of generated samples $G(\mathbf{z})$ obtained from passing 
latent vectors $\mathbf{z}$ through $G$ such that $\mathbf{z} \sim p_{\mathbf{z}}$. The learning of the generator 
means that the samples evolve over the training process. Eventually, the samples will appear to have been drawn 
from the true data distribution $p_{data}$ so that $G(\mathbf{z}) \sim p_{data}$. The generated samples 
$\mathbf{x} \sim p_{model}$ and real samples $\mathbf{x} \sim p_{data}$ become indistinguishable.

@@

During training the discriminator minimises $J^{(D)}$ only through feedback from backpropagation to its parameters 
$\bm{\theta}^{(D)}$. The generator $G$ minimises $J^{(G)}$ by only affecting its parameters $\bm{\theta}^{(G)}$ through 
returns from backpropagation related to the discriminator classifications. The minimization objectives are simultaneous 
between two networks in search of a Nash equilibrium and added together the objectives for the discriminator and generator 
give a value function $V(\bm{\theta}^{(D)}, \bm{\theta}^{(G)})$ in the combined parameter space $(\bm{\theta}^{(D)}, 
\bm{\theta}^{(G)})$. Figure~\ref{fig:gan_game} shows the value function in the combined parameter space for which the 
backpropagation gradients are calculated with $D$ and $G$ as adversaries of each other. The optimum parameters for each 
network are shown at the Nash equilibrium. The game between the models is described with the value function $V$ as

@@

in the combined parameter space. This equation shows the opposition between the maximization of the correct 
classifications by $D$ and the minimisation of the correct classifications by $D$ of the generator products that are 
made by $G$.

@@

this is from the cross-entropy discriminator objective function for the binary classification of real and fake 
samples. Appendix~\ref{appendix:cross_entropy} defines cross-entropy in this context. The expectation operators 
$\mathbb{E}_{\mathbf{x} \sim p_{data}}$ and $\mathbb{E}_{G(\mathbf{z}) \sim p_{model}}$ correspond to the ensemble 
averages over the classifications of real and fake samples respectively. Equation~\ref{eq:J_D} shows that the first 
term cannot be influenced by the generator $G$ because here the samples are drawn from the data distribution to which 
it has no access to. Inputs to the generator $\mathbf{z}$ are drawn from the prior distribution $p_\mathbf{z}$ over 
the latent-space to the model. The discriminator is given a generator sample $G(\mathbf{z})$ to classify it as fake, 
so $D(G(\mathbf{z})) \! \rightarrow \! 0$ whereas the generator $G$ tries to make $D(G(\mathbf{z})) \!  \rightarrow \! 1$. 
The Nash equilibrium is satisfied for the model and data distributions becoming identical (in the ideal case) so that 
$G(\mathbf{z})$ is drawn from $p_{data}$. This means the discriminator classification tends to $D(\mathbf{x}) = 
\frac{1}{2}$ for all samples~$\mathbf{x}$~\cite{gf_gan}. The optimum of the generator $G$ parameters 
$\bm{\theta}^{(G)*}$ is

@@

Rodriguez et al. 2018~\cite{web_gan} use GANs to generate realistic 2D samples of N-body simulation mass distributions 
of size $100 \: h^{-1} \,\textup{Mpc}$ and $500 \: h^{-1} \,\textup{Mpc}$. This paper is the closest to the work 
presented in this report. N-body simulation data was obtained from the particle distribution of the L-PICOLA 
simulation~\cite{LPICOLA} at $z=0$. This simulation is based on the $\Lambda \text{CDM}$ cosmological model with 
the Hubble constant $H_0=70\textup{ km s}^{-1}\textup{ Mpc}^{-1}$, dark energy density $\Omega_\Lambda=0.72$ and
matter density $\Omega_m=0.28$. The method of Rodriguez et al. 2018 differs from this report in the statistical 
tests of the generator output. The generated samples are cross-correlated with each other in pairs to quantify 
their mutual independence. This estimates the effect of a weakness of GANs against other generative models, known
as mode collapse~\cite{gf_gan}, where generated samples can often be very similar if not identical. The generated 
2D slices are compared to simulation samples in a cosmological context by calculating the automatic and cross 
(sample-to-sample) power spectrum of the 2D images using a discrete Fourier transform. The power spectra $P_k$ 
(see Appendix~\ref{appendix:power_spectrum}) of the generated samples against the training samples agree to 1-2\% 
for the majority of $k$ values. This work shows the potential for GAN methods to subvert N-body simulations.

@@

Minibatch discrimination~\cite{gantricks_sali} was implemented to help avoid mode collapse by discriminating over a 
collection of samples given to the discriminator as opposed to a single sample at a time. The original work describes 
the `closeness' of samples in a collection as follows: Let $\mathbf{f}(\bm{x}_i) \in \mathbb{R}^A$ represent a vector of 
features extracted by $D$ for sample $\mathbf{x}_i$. Multiplying $\mathbf{f}(\bm{x}_i)$ by a tensor $T \in \mathbb{R}^{A\times 
B\times C}$ results in a matrix $M_i^{B\times C}$. The $L_1$-distance is calculated between each row of $M_i$ for each sample 
$i$ in the collection. The distances are exponentiated as

@@

Where $\mathbf{X}$ denotes a minibatch of samples. The minibatch discrimination output is concatenated to the 
flattened-features from the last convolutional layers in the discriminator. This is done separately for the real and 
fake samples. The discriminator is able to designate the validity of a sample with additional information from samples 
in a collection.

@@

\begin{figure}
\includegraphics[width=\columnwidth]{figures/diagrams/kl_stat_process3.png}
\centering
\caption{From a single ensemble of 3D-histograms to a normalised distribution; 
         (a) - an ensemble of 3D-histograms that will be transformed into a distribution, 
         (b) - the flattened vector of bin-cells, 
         (c) - the histograms of each vector-component, 
         (d) - the final collapsed vector encoding the whole of a given distribution of 3D-histograms.}
\label{fig:klstat}
\end{figure}

@@

Figure~\ref{fig:acgan_interp} shows the results of varying the class label for a single latent-space point. Between the 
minimum and maximum redshifts $z=0.0$ and $z=5.0$ respectively, 40 class examples were generated.

@@

\subsubsection{Latent space point interpolation}

@@

The new redshifts are $z'=\{0.51, 1.50, 2.62, 3.58, 4.52\}$ where the original redshift values $z$ shown to the
ACGAN are $z=\{0.0, 1.07, 2.07, 3.06, 4.17, 5.28\}$. With these new redshifts the same $D_{KL}$-trials were repeated and 
the results are shown in 
% plot unseen z cubes against true redshifts: 5 rows, 6 cols: 3 real, 3 gen

@@

% LOSSES AND ACCURACIES
Figure~\ref{fig:acgan_losscurve} shows the ACGAN objectives and accuracies over a run of training. The behaviours of the 
models over the training are apparent in these curves. The sharp jump at the start over the first few $10^3$ epochs or so 
shows the initial learning for both models that are new to their tasks. Each sharp jump in the losses of both models shows
the  generator attempting a new strategy for its objective. There are small fluctuations in the accuracies of either model 
as the generator makes smaller adjustments relative to the initial training, just after $1.0\times10^4$ epochs, these are 
in part due to the mistakes made by the discriminator in seeing new training and generated samples.

The accuracies reported by the discriminator model tend very slowly toward around $55\%$. This is an ideal value because the
discriminator and generator are both engaged in a conflict that still supplies information through gradients to the generator
model. The samples are still being improved on at this stage. This is echoed in the convergence of the losses for each model 
to a value of around $9.25$. 
This does not mean an ideal `perfect' generator has closed the distance between the model and data distributions, but instead 
that the architecture of the discriminator cannot extract any more useful information (and is extracting less and less) to 
the generator.The gradients for backpropagation have vanished. This was seen in training. The generator suffered from mode 
collapse and made more samples with more similarity between them.

The accuracies of both models converge at just over 60\% which implies there is space to improve the model 
architectures. The separate networks are not able to fully attain their objectives against each other. At this point the 
discriminator is extracting the maximal amount of information for use in classification and feedback to the generator, 
which was observed to be changing the histograms only at the scales given by its convolutional kernels. In other words, 
the generator is fooling the discriminator without generating the most realistic cubes it could. Through running many 
different models it was found that $8.0\times 10^4$ epochs was enough time for the network to converge. Past this point 
the training would dissipate chaotically.

@@

% KL DIVERGENCES
The KL divergences between the training and generated distributions $t$ and $g$, denoted by $D_{KL}(t||g)$ and 
$D_{KL}(g||t)$, follow a path that is related to the behaviour of the models shown in loss and accuracy curve. The 
divergences both start at a global maximum and tend to a final convergent value. The initial value corresponds to the 
comparison between batches of cubes with pixel-densities sampled from the $\mathbf{z}$-prior $\mathcal{N}(0,1)$ that are
devoid of structure and the final value corresponds to the cubes with realistic cosmological structure in. The values 
sharply drop at the same epoch in training as the losses and accuracies do. The initial $D_{KL}$ values of the comparison 
are with earlier generated images that are devoid of structure. The final value, with some intermediate values, corresponds
to images that are very similar to the training data, so the $D_{KL}$ values changed over a range of $10^{1.0}$ to 
$10^{-1.4}$ from structureless images to realistic images.

The curves have similar features to the loss and accuracy curves as they should because of the relation of the 
cross-entropy to the KL divergence $D_{KL}$.

The increasingly small differences between the observed $D_{KL}(t||g)$ and $D_{KL}(g||t)$ values with respect to the values 
throughout training is important. This suggests that the training and generated distributions $t$ and $g$ are suited for 
replacement with each other, as opposed to a preferred replacement of one with the other.

@@

Figure~\ref{fig:kl_trials_all} and Table~\ref{tab:1} show the results of the $D_{KL}$-trials for batches of 32 samples from 
each distribution repeated for 64 separate comparisons between each distribution pairing. The procedure for the repeated 
ensemble KL divergence testing outlined in Section~\ref{methods:stats_and_diags} used the samples transformed into 
probability distributions over the pixel density.

There is a slight broadening of the histograms with increasing kernel size. This shows the random effect, or dispersion, 
on the $D_{KL}$ value from repeated tests to on separate ensembles of histograms. This is caused by the loss of information 
due to the down-sampling of larger areas causing more variance in a given value of pixel density because of the averaging 
over the original pixels. This is predicted by the central-limit theorem. In this discussion a `mutual comparison' 
refers to the KL divergence $D_{KL}$ between two different ensembles from the same distribution of $t$, $g$ or $u$. The 
mutual $D_{KL}$ comparisons that show the most broadening are the mutual comparisons which together have increasing 
variance and separation with kernel size. This shows that in both cases the distributions are composed very differently 
to the uniform distributions. The means of the mutual distributions stay the same except for the uniform-uniform comparison. 
This is because the median for the bin separations in the component-histograms was given by the median of the training data 
batch which consistently has a lower mean pixel density than the uniform histogram data.

\begin{figure*}[hbt!]
\includegraphics[width=17cm]{figures/graphs/kl_trials_all_z_train1.png}
\centering
\caption{Histograms showing the $D_{KL}$ values for distribution-comparisons between training $t$, generated $g$ and 
         uniform $u$ cube samples. This graph shows the results for downsampling kernel values $k \in \{1, 2, 4\}$.}
\label{fig:kl_trials_all}
\end{figure*}

@@

The pixel density distribution is an important characteristic of the images but the just obtaining the 
similarity between the $t$ and $g$ distributions with this property would not guarantee that all the structure in the 
training data was present. There are many combinations of pixels that make up a matching pixel-density distribution for 
two histogram distributions, but far less of those show the structure of the cosmic web. 
% The higher standard deviations in the $D_{KL}$ histograms for the 2D case in all the kernel sizes compared to the 3D cases is due to the larger size of features relative to total size of the image that shows them in either case.

@@

Figure~\ref{fig:kl_trials_all_z2} and Table~\ref{tab:2} show the results of the $D_{KL}$-trials between for batches of 32 
samples from each distribution repeated for 64 separate comparisons between each distribution pairing. These results are 
produced by using the generator to create samples at values of redshift $z=z'$ that the ACGAN has never been exposed to.

@@

The results point to the fact that the generator is less competent at representing values of redshift that lay between the
values that the ACGAN has been exposed to in training. This is certainly to be expected as the description for the emergence
of the finite training set of images is certainly more simple than the chronology of the cosmic web that the Millenium-SAGE
models characterise. Figure~\ref{fig:z_kl} shows the relation between the $\langle D_{KL}(g||t) \rangle$ and 
$\langle D_{KL}(t||g) \rangle$ values with redshift $z$. The highest peak coincides with the lowest unseen redshift $z'=0.51$.

The last two unseen redshift values show consistency between kernel sizes and they are similar values to the trained redshift 
values either side of these redshifts. Because these $\langle D_KL \rangle$ values have the lowest separation between the 
nearest trained redshifts, it suggests that the ACGAN has a certain range in redshift over which it can obtain the same coherence 
in samples with. The other redshifts suggest this boundary away from a given trained redshift value is between 0.3 and 0.5.

@@

\subsection{Improvements to the present method}
The strongest dependence of the quality and realism of the generator product are the GAN models and hyper-parameters. It 
should be noted that each of these should really be tested independently but it is quite possible that a lifetime is not 
long enough to do so. The author's opinion is that the best route is to improve the current architecture by increasing the 
number of pixels in the cubes, allowing the depth of the models to increase with more layers, whilst also better translating
the variance in the structure inside the histograms. It has been found that in amplifying the initial parameterised units 
in the first layer of the generator model requires that there are at least $\sim \! 10^4$ units. The use of
fewer units does not allow the generator to draw enough permutations from the potentials  `proto-cubes' allowed by the 
initial layer. The kernels in the convolutional layers were found to be most effective at size $9 \times 9 \times 9$. 
Though it is expected that some combination of higher kernel sizes would better characterise the histogram structures.

@@

The other available galaxy properties from the Millennium-SAGE catalogs could also be folded into the generator analysis
to create distributions weighted by luminosity or another characteristic. It is possible that some amount of control over
the generated images and their additional characteristics

In previous work, a regular GAN was trained on the same $z=0$ data and a smaller difference in the mean $D_{KL}(t||g)$ value
was found. As a one class model, this suggests that the architectures in this ACGAN implementation could require a higher 
capacity in terms of convolution filters.
% 3D model works REALLY well - use transposed conv instead of upsampling. As a one-class model, the capacity implies that the
% ACGAN requires more filters that the GPU could not provide.

@@

\subsection{Universe generation} 
To create a cosmological volume on order of the size of the Millennium-SAGE volume, with the tested default cube-size, a 
generator would need to create around $10^2$ cubes. The generator model in this work once trained can do so within 30 
seconds. Whilst in the process of fusing the samples together coherently avoiding density discontinuities would not be a 
trivial task. It is logical to assume that it is a process that is maximally as intensive as generating the cubes 
themselves. It is expected that the GAN would be able to generalise well to higher resolution or larger volume sample data.

>>>>

\begin{figure}
\includegraphics[width=\columnwidth, trim=4 4 4 4,clip]{figures/diagrams/ACGAN_diagram_box.png}
\centering
\caption{Schematic of the GAN framework; including discriminator and generator networks, the latent-space vector, dataset 
         of samples and the backpropagation feedback to model parameters. \textit{Source} and \textit{Class} refer to the
         values of $p_\mathbf{x}$ and $p_\mathbf{c}$ designated by the discriminator on a sample $\mathbf{x}$.}
\label{fig:GAN_diagram}
\end{figure}

@@

% WHAT ARE D AND G? HOW DO D AND G WORK?
% The discriminator and generator models can be represented as functions $D$ and $G$ respectively because of their 
% mappings as neural networks and parameter dependencies. 
The discriminator $D$ has image input $\mathbf{x}$ and parameters 
$\bm{\theta}^{(D)}$. The generator $G$ takes a latent space vector $\mathbf{z}$ as input and uses parameters 
$\bm{\theta}^{(G)}$. 
% Information flows through the network and its parameters to create the outputs $D(\mathbf{x})$ or 
% $D(G(\mathbf{z}))$ for the discriminator and $G(\mathbf{z})$ for the generator. This is the forward-propagation from 
% an input to an output. The model performances are calculated with the objective functions $J^{(D)}$ or $J^{(G)}$.

% HOW DO D AND G WORK TOGETHER IN A GAN AND PLAY THE GAME?
In the case of a GAN, the optimisation landscape is more complex because of the simultaneous learning of the two 
neural networks that both constantly adjust their parameters. 
% The discriminator and generator models in a GAN depend on parameters $\bm{\theta}^{(D)}$ and $\bm{\theta}^{(G)}$ respectively. 
% The models typically consist of multi-layer 
% perceptrons with the potential for other parametric components such as convolutional layers and the parameters are 
% the state of these components at an epoch of training. 
The generator $G$ draws noise vectors $\mathbf{z}$ from a prior 
distribution $p_{\mathbf{z}}$. The vectors $\mathbf{z}$ are passed through the generator architecture to create a 
sample $\mathbf{x}=G(\mathbf{z};\bm{\theta}^{(G)})$ that is now a sample $\mathbf{x}$ from the model distribution 
$p_{model}$ - this is the distribution of generated samples $G(\mathbf{z})$ obtained from passing 
latent vectors $\mathbf{z}$ through $G$ such that $\mathbf{z} \sim p_{\mathbf{z}}$. The learning of the generator 
means that the samples evolve over the training process. Eventually, the samples will appear to have been drawn 
from the true data distribution $p_{data}$ so that $G(\mathbf{z}) \sim p_{data}$. The generated samples 
$\mathbf{x} \sim p_{model}$ and real samples $\mathbf{x} \sim p_{data}$ become indistinguishable.

@@

During training the discriminator minimises $J^{(D)}$ only through feedback from backpropagation to its parameters 
$\bm{\theta}^{(D)}$. The generator $G$ minimises $J^{(G)}$ by only affecting its parameters $\bm{\theta}^{(G)}$ through 
returns from backpropagation related to the discriminator classifications. The minimization objectives are simultaneous 
between two networks in search of a Nash equilibrium and added together the objectives for the discriminator and generator 
give a value function $V(\bm{\theta}^{(D)}, \bm{\theta}^{(G)})$ in the combined parameter space $(\bm{\theta}^{(D)}, 
\bm{\theta}^{(G)})$. Figure~\ref{fig:gan_game} shows the value function in the combined parameter space for which the 
backpropagation gradients are calculated 
%with $D$ and $G$ as adversaries of each other. 
The optimum parameters for each 
network are shown at the Nash equilibrium. The game between the models is described with the value function $V$ as

@@

in the combined parameter space. The expectation operators 
$\mathbb{E}_{\mathbf{x} \sim p_{data}}$ and $\mathbb{E}_{G(\mathbf{z}) \sim p_{model}}$ correspond to the ensemble 
averages over the classifications of real and fake samples respectively. This equation shows the opposition between 
the maximization of the correct classifications by $D$ and the minimisation of the correct classifications by $D$ of 
the generator products that are made by $G$.

@@

this is from the cross-entropy discriminator objective function for the binary classification of real and fake 
samples. Appendix~\ref{appendix:cross_entropy} defines cross-entropy in this context. Equation~\ref{eq:J_D} shows that the first 
term cannot be influenced by the generator $G$ because here the samples are drawn from the data distribution to which 
it has no access to. Inputs to the generator $\mathbf{z}$ are drawn from the prior distribution $p_\mathbf{z}$ over 
the latent-space to the model. The discriminator is given a generator sample $G(\mathbf{z})$ to classify it as fake, 
so $D(G(\mathbf{z})) \! \rightarrow \! 0$ whereas the generator $G$ tries to make $D(G(\mathbf{z})) \!  \rightarrow \! 1$. 
The Nash equilibrium is satisfied for the model and data distributions becoming identical (in the ideal case) so that 
$G(\mathbf{z})$ is drawn from $p_{data}$. This means the discriminator classification tends to $D(\mathbf{x}) = 
\frac{1}{2}$ for all samples~$\mathbf{x}$~\cite{gf_gan}. The optimum of the generator $G$ parameters 
$\bm{\theta}^{(G)*}$ is

@@

Rodriguez et al. 2018~\cite{web_gan} use GANs to generate realistic 2D samples of N-body simulation mass distributions 
of size $100 \: h^{-1} \,\textup{Mpc}$ and $500 \: h^{-1} \,\textup{Mpc}$. This paper is the closest to the work 
presented in this report. N-body simulation data was obtained from the particle distribution of the L-PICOLA 
simulation~\cite{LPICOLA} at $z=0$. This simulation is based on the $\Lambda \text{CDM}$ cosmological model with 
the Hubble constant $H_0=70\textup{ km s}^{-1}\textup{ Mpc}^{-1}$, dark energy density $\Omega_\Lambda=0.72$ and
matter density $\Omega_m=0.28$. The method of Rodriguez et al. 2018 differs from this report in the statistical 
tests of the generator output. The generated samples are cross-correlated with each other in pairs to quantify 
their mutual independence. This estimates the effect of a weakness of GANs against other generative models, known
as mode collapse~\cite{gf_gan}, where generated samples can often be very similar if not identical. The generated 
2D slices are compared to simulation samples in a cosmological context by calculating the automatic and cross 
(sample-to-sample) power spectrum of the 2D images using a discrete Fourier transform. The power spectra $P_k$ 
% (see Appendix~\ref{appendix:power_spectrum}) 
of the generated samples against the training samples agree to 1-2\% 
for the majority of $k$ values. This work shows the potential for GAN methods to subvert N-body simulations.

@@

\subsubsection{Minibatch Discrimination}

Minibatch discrimination~\cite{gantricks_sali} was implemented to help avoid mode collapse by discriminating over a 
collection of samples given to the discriminator as opposed to a single sample at a time. The original work describes 
the `closeness' of samples in a collection as follows: Let $\mathbf{f}(\bm{x}_i) \in \mathbb{R}^A$ represent a vector of 
features extracted by $D$ for sample $\mathbf{x}_i$, where $A$ is the number of components of the feature vector. Multiplying $\mathbf{f}(\bm{x}_i)$ by a tensor $T \in \mathbb{R}^{A\times B\times C}$ results in a matrix $M_i^{B\times C}$, where $B$ 
and $C$ are the number of discriminative kernels and the dimensionality of the space where the closeness of samples is 
calculated. The $L_1$-distance is calculated between each row of $M_i$ for each sample $i$ in the collection. The distances 
are exponentiated as

@@

Where $n$ is the number of pairings in a minibatch and  $\mathbf{X}$ denotes a minibatch of samples. The minibatch 
discrimination output is concatenated to the 
flattened-features from the last convolutional layers in the discriminator. This is done separately for the real and 
fake samples. The discriminator is able to designate the validity of a sample with additional information from samples 
in a collection which in turn causes the generator to promote variety in its products.

@@

\begin{figure}
\includegraphics[width=\columnwidth, trim=4 4 4 4,clip]{figures/diagrams/kl_stat_process1.png}
\centering
\caption{From a single ensemble of 3D-histograms to a normalised distribution; 
         (a) - an ensemble of 3D-histograms that will be transformed into a distribution, 
         (b) - the flattened vector of bin-cells, 
         (c) - the histograms of each vector-component, 
         (d) - the final collapsed vector encoding the whole of a given distribution of 3D-histograms.}
\label{fig:klstat}
\end{figure}

@@

Figure~\ref{fig:acgan_interp} shows the results of varying the class label for a single latent-space point. Between the 
minimum and maximum redshifts $z=0.0$ and $z=5.28$ respectively, 40 class examples were generated. The figure shows a 
smooth evolution of the latent space point evaluated by the generator for redshift $z$. It is left to Section~\ref{sec:kl_trials} 
for a discussion on the validity of the generator samples that are conditioned for $z$ values not in the set of training 
data redshifts, though initial inspection suggests the generator is able to reproduce samples of all redshifts consistently.

@@

% \subsubsection{Latent space point interpolation}

@@

The new values of simulation snapshot redshifts are $z'=\{0.51, 1.50, 2.62, 3.58, 4.52\}$ where the original redshift values
$z$ shown to the ACGAN are $z~=~\{0.0, 1.07, 2.07, 3.06, 4.17, 5.28\}$. With these new redshifts the same $D_{KL}$-trials 
were repeated and the results are shown in \ref{fig:kl_trials_all_z2}. A number of similar redshift values to the unseen
redshift values $z'$ are displayed in Figure~\ref{fig:acgan_interp}.
% plot unseen z cubes against true redshifts: 5 rows, 6 cols: 3 real, 3 gen

@@

% LOSSES AND ACCURACIES
Figure~\ref{fig:acgan_losscurve} shows the ACGAN objectives and accuracies over a run of training. The behaviours of the 
models over the training are apparent in these curves. The initial decline over the first few $10^4$ epochs or so 
shows the initial learning for both models that are new to their tasks. Each sharp jump in the losses of both models shows
the  generator attempting a new strategy for its objective. There are small fluctuations in the accuracies of either model 
as the generator makes smaller adjustments relative to the initial training, just after $1.0\times10^4$ epochs, these are 
in part due to the mistakes made by the discriminator in seeing new training and generated samples. 
% The use of spectral normalisation keeps the weights of the % general

The accuracies and losses of the models in this implementation are extremely chaotic in comparison to typical behaviours of
these parameters in GANs. This is because the spectral normalization regularization constrains the spectral norm of the weight
matrices on each layer it is applied to. This means that changes to this value are always normalised in the next training
iteration. This chaotic behaviour is implicit to the GAN framework and not this implementation itself. The cross entropy loss
is related to the specific KL-divergence we have defined in this work and so the models converging can still be seen in the
$\langle D_{KL}\rangle_z$ curves.

% The accuracies reported by the discriminator model tend very slowly toward around $55\%$. This is an ideal value because the
% discriminator and generator are both engaged in a conflict that still supplies information through gradients to the generator
% model. The samples are still being improved on at this stage. This is echoed in the convergence of the losses for each model 
% to a value of around $9.25$. 
% This does not mean an ideal `perfect' generator has closed the distance between the model and data distributions, but instead 
% that the architecture of the discriminator cannot extract any more useful information (and is extracting less and less) to 
% the generator.The gradients for backpropagation have vanished. This was seen in training. The generator suffered from mode 
% collapse and made more samples with more similarity between them. % NON SN model

% The accuracies of both models converge at just over 60\% which implies there is space to improve the model 
% architectures. The separate networks are not able to fully attain their objectives against each other. At this point the 
% discriminator is extracting the maximal amount of information for use in classification and feedback to the generator, 
% which was observed to be changing the histograms only at the scales given by its convolutional kernels. In other words, 
% the generator is fooling the discriminator without generating the most realistic cubes it could. Through running many 
% different models it was found that $8.0\times 10^4$ epochs was enough time for the network to converge. Past this point 
% the training would dissipate chaotically.

@@

The KL divergences between the training and generated distributions $t$ and $g$, denoted by $\langle D_{KL}(t||g) \rangle_z$ 
and $\langle D_{KL}(g||t) \rangle_z$ (where $\langle\dot\rangle_z$ is an averaging over redshift), follow a path that is 
related to the behaviour of the models shown in loss and accuracy curve. The divergences both start at a global maximum and 
tend to a final convergent value. The initial value corresponds to the comparison between batches of cubes with 
pixel-densities sampled from the $\mathbf{z}$-prior $\mathcal{N}(0,1)$ that are devoid of structure and the final value 
corresponds to the cubes with realistic cosmological structure in. 
% The values sharply drop at the same epoch in training as the losses and accuracies do. 
The initial $D_{KL}$ values of the comparison are with earlier generated images that are 
devoid of structure. The final value, with some intermediate values, corresponds to images that are very similar to the 
training data, so the $D_{KL}$ values changed over a range of $10^{1.0}$ to $10^{-1.4}$ from structureless images to 
realistic images.

% The curves have similar features to the loss and accuracy curves as they should because of the relation of the 
% cross-entropy to the KL divergence $D_{KL}$.

The increasingly small differences between the observed $\langle D_{KL}(t||g) \rangle_z$ and $\langle D_{KL}(g||t) \rangle_z$ 
values with respect to the values throughout training is important. This suggests that the training and generated distributions 
$t$ and $g$, and ultimately $p_{model}$ and $p_{data}$ are suited for replacement with each other, as opposed to a preferred 
replacement of one with the other.

@@

Figure~\ref{fig:kl_trials_all_z1} shows the results of the $D_{KL}$-trials for batches of 64 samples comprised from samples
from each distribution repeated for 128 separate comparisons between each distribution pairing. The procedure for the 
repeated ensemble KL divergence testing outlined in Section~\ref{methods:stats_and_diags} used the samples transformed into 
probability distributions over the pixel density.

There is a slight broadening of the histograms with increasing kernel size. This shows the random effect, or dispersion, 
on the $D_{KL}$ value from repeated tests to on separate ensembles of histograms. This is caused by the loss of information 
due to the down-sampling of larger areas causing more variance in a given value of pixel density because of the averaging 
over the original pixels. This is predicted by the central-limit theorem. In this discussion a `mutual comparison' 
refers to the KL divergence $D_{KL}$ between two different ensembles from the same distribution of $t$, $g$ or $u$. The 
mutual $D_{KL}$ comparisons that show the most broadening are the mutual comparisons which together have increasing 
variance and separation with kernel size. This shows that in both cases the distributions are composed very differently 
to the uniform distributions. The means of the mutual distributions stay the same except for the uniform-uniform comparison. 
This is because the median for the bin separations in the component-histograms was given by the median of the training data 
batch which consistently has a lower mean pixel density than the uniform histogram data. 
% It is also not symmetric as can be seen from the sheer sparsity of the samples for each redshift.

\begin{figure*}[hbt!]
\includegraphics[width=17cm]{figures/graphs/kl_trials_all_z_train1.png}
\centering
\caption{Histograms showing the $D_{KL}$ values for distribution-comparisons between training $t$, generated $g$ and 
         uniform $u$ cube samples. This graph shows the results for downsampling kernel values $k \in \{1, 2, 4\}$.}
\label{fig:kl_trials_all_z1}
\end{figure*}

@@

The pixel density distribution is an important characteristic of the images but just obtaining the 
similarity between the $t$ and $g$ distributions with this property would not guarantee that all the structure in the 
training data was present. There are many combinations of pixels that make up a matching pixel-density distribution for 
two histogram distributions, but far less of those show the structure of the cosmic web. 
% The higher standard deviations in the $D_{KL}$ histograms for the 2D case in all the kernel sizes compared to the 3D cases is due to the larger size of features relative to total size of the image that shows them in either case.

@@

Figure~\ref{fig:kl_trials_all_z2} shows the results of the $D_{KL}$-trials between for batches of 64 
samples from each distribution repeated for 128 separate comparisons between each distribution pairing. These results are 
produced by using the generator to create samples at values of redshift $z=z'$ that the ACGAN has never been exposed to.

@@

The results point to the fact that the generator is competent at representing values of redshift that lay between the
values that the ACGAN has been exposed to in training. This is interesting because the description for the emergence
of the finite training set of images is certainly more simple than the chronology of the cosmic web that the Millenium-SAGE
models characterise. Figure~\ref{fig:z_kl} shows the relation between the $\langle D_{KL}(g||t) \rangle$ and 
$\langle D_{KL}(t||g) \rangle$ values with redshift $z$. The highest peak coincides with the lowest unseen redshift $z'=0.51$.

The last two unseen redshift values show consistency between kernel sizes and they are similar values to the trained redshift 
values either side of these redshifts. Because these $\langle D_{KL} \rangle$ values have the lowest separation between the 
nearest trained redshifts, it suggests that the ACGAN has trouble obtaining the same coherence in samples either side of the 
median $z$ value of the data it is exposed to. The coherence of the $z$ redshift samples to the $z'$ redshift samples suggests
the ACGAN can represent samples to a distance away from a given trained redshift value of between 0.3 and 0.5.

The similarity between the lowest and highest $z$ and $z'$ $D_{KL}$-histograms suggests that the characteristics of the data 
cause the ACGAN some difficulty in imitating samples at these redshifts successfully, but that it is not necessarily a deficiency 
that only this implementation of the ACGAN would suffer with.

@@

\subsection{Improvements to the present method}
The strongest dependence of the quality and realism of the generator product are the GAN models and hyper-parameters. It 
should be noted that each of these should really be tested independently but it is quite possible that a lifetime is not 
long enough to do so. The author's opinion is that the best route is to improve the current architecture by increasing the 
number of pixels in the cubes, allowing the depth of the models to increase with more layers, whilst also better translating
the variance in the structure inside the histograms. It has been found that in amplifying the initial parameterised units 
in the first layer of the generator model requires that there are at least $\sim \! 10^4$ units. The use of
fewer units does not allow the generator to draw enough permutations from the potential `proto-cubes' allowed by the 
initial layer. The kernels in the convolutional layers were found to be most effective at size $9 \times 9 \times 9$. 
Though it is expected that some combination of higher kernel sizes would better characterise the histogram structures.

@@

The other available galaxy properties from the Millennium-SAGE catalogs could also be folded into the generator analysis
to create distributions weighted by luminosity or another characteristic. It is possible that some amount of control over
the generated images and their additional characteristics could deepen the effectiveness of replacing simulation data in 
analyses, allowing samples that are emergent from different cosmological parameters to be considered separately.

In previous work, a regular GAN was trained on the same $z=0$ data and a smaller difference in the mean $D_{KL}(t||g)$ value
was found. As a one class model, this suggests that the architectures in this ACGAN implementation could require a higher 
capacity in terms of convolution filters. The differences in the pixel sparsity between redshift classes increases the need
for more filters in the ACGAN convolutions.
% 3D model works REALLY well - use transposed conv instead of upsampling. As a one-class model, the capacity implies that the
% ACGAN requires more filters that the GPU could not provide.

@@

\subsection{Universe generation} 
To create a cosmological volume on order of the size of the Millennium-SAGE volume, with the tested default cube-size, a 
generator would need to create around $20$ cubes. The generator model in this work once trained can do so within 
seconds. Whilst in the process of fusing the samples together coherently avoiding density discontinuities would not be a 
trivial task, it is logical to assume that it is a process that is maximally as intensive as generating the cubes 
themselves. It is expected that the GAN would be able to generalise well to higher resolution or larger volume sample data.
